---
title: "Oasis Scourge Documentation"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
    css: "../site/styles/custom_styles.css"
knit: (function(inputFile, encoding) {
    rmarkdown::render(inputFile, encoding = encoding, output_dir = "../site")
  })
---

```{r setup, include=FALSE}
library(flexdashboard)
```

<!-- Complete Integrated Javascript Logic Engine -->
```{r, echo=FALSE, results='asis'}
cat('
<script type="text/javascript">
document.addEventListener("DOMContentLoaded", function() {
  const root = document.documentElement;
  const btn = document.getElementById("themeBtn");
  
  // 1. Restore Dark Mode Theme Token
  const savedTheme = localStorage.getItem("dashboardTheme");
  if (savedTheme === "dark") {
    root.setAttribute("data-theme", "dark");
    if (btn) btn.innerHTML = "🌙 Dark Mode";
  } else {
    root.removeAttribute("data-theme");
    if (btn) btn.innerHTML = "☀️ Light Mode";
  }

  // 2. FIX STATE RECOVERY: Force sidebar link highlighting to match the active URL hash on refresh
  const currentHash = window.location.hash;
  if (currentHash) {
    // Strip default active highlights
    document.querySelectorAll(".custom-nav-link").forEach(link => link.classList.remove("active"));
    
    // Find the link matching this precise browser hash and illuminate it
    const activeLink = document.querySelector(`.custom-nav-link[href="${currentHash}"]`);
    if (activeLink) {
      activeLink.classList.add("active");
    }
  }
});

function toggleDashboardTheme() {
  const root = document.documentElement;
  const btn = document.getElementById("themeBtn");
  const currentTheme = root.getAttribute("data-theme");
  
  if (currentTheme === "dark") {
    root.removeAttribute("data-theme");
    btn.innerHTML = "☀️ Light Mode";
    localStorage.setItem("dashboardTheme", "light");
  } else {
    root.setAttribute("data-theme", "dark");
    btn.innerHTML = "🌙 Dark Mode";
    localStorage.setItem("dashboardTheme", "dark");
  }
}

function styleSelectedTab(element) {
  document.querySelectorAll(".custom-nav-link").forEach(link => link.classList.remove("active"));
  element.classList.add("active");
}
</script>
')
```

<!-- Dark Mode Interface Controller -->
<div class="theme-switch-wrapper">
  <button class="theme-switch-btn" id="themeBtn" onclick="toggleDashboardTheme()">🌙 Dark Mode</button>
</div>


Global Navigation {.sidebar data-width=400}
=====================================

### Navigation Options

This persistent sidebar stays pinned to the left across all pages. The links above instantly switch views via native CSS without any flickering.

Below are tabs for the Oasis Scourge and Wyvern's Mark regional cultures.   A regional culture is a unique culture defined not merely by regions, but whether enough material exists on the race within that region to warrant documentation here.   If a regional-race is not included, it does not mean irrelevance, but rather the game environment will disclose the cultural dynamics of the folk residing there.  

<br>
<hr>
<br>


<ul class="custom-nav-list">
  <li><a href="#intro" class="custom-nav-link active" onclick="styleSelectedTab(this)">Introduction</a></li>
  <li><a href="#hobbog-salt-willow" class="custom-nav-link" onclick="styleSelectedTab(this)">HobBogs of Salt-Willow</a></li>
  <li><a href="#hobbog-duskreed" class="custom-nav-link" onclick="styleSelectedTab(this)">HobBogs of Dusk Reed</a></li>
  <li><a href="#cragwold" class="custom-nav-link" onclick="styleSelectedTab(this)">Cragwolds of the Weald</a></li>
  
  <li><a href="#glenkin-dellari" class="custom-nav-link" onclick="styleSelectedTab(this)">Glenkin: The Delari Diaspora</a></li>
  <li><a href="#glenkin-ashvari" class="custom-nav-link" onclick="styleSelectedTab(this)">Glenkin: The Ashvari of Glentali</a></li>
  <li><a href="#songs-salt-willow" class="custom-nav-link" onclick="styleSelectedTab(this)">The Songs of Salt-Willow</a></li>
</ul>


Intro {#intro .hidden}
=====================================

```{r, child='01-intro.Rmd'}
```


HobBogs_Salt_Willow {#hobbog-salt-willow .hidden}
=====================================

```{r, child='02-hobbog-salt-willow.Rmd'}
```



HobBogs_Dusk_Reed {#hobbog-duskreed .hidden}
=====================================

```{r, child='03-hobbog-duskreed.Rmd'}
```



Cragwolds {#cragwold .hidden}
=====================================

```{r, child='04-cragwold.Rmd'}
```






Glenkin_Dellari {#glenkin-dellari .hidden}
=====================================

```{r, child='05-glenkin_dellari.Rmd'}
```







Glenkin_Ashvari {#glenkin-ashvari .hidden}
=====================================

```{r, child='06-glenkin_ashvari.Rmd'}
```




<br>
<br>
<br>
<hr>
<br>

#### Songs of Oasis Scourge

Glenkin_Ashvari {#songs-salt-willow .hidden}
=====================================

```{r, child='20-songs-hobbogs-saltwillow.Rmd'}
```





